\input{111pre}
\begin{document}
\MYTITLE{Lab 6 \\ 30 September 2015 \\ Due Wednesday, 7 October by 2:30pm }

\subsection*{Objectives}
\vspace{-0.05in}

In addition to enhancing your teamwork skills, to write a Java program that manipulates strings of DNA by appropriately
using methods from the {\tt java.util.String} and {\tt java.util.Random} classes.

\vspace{-0.05in}
\subsection*{Important Notes}
\vspace{-0.05in}

This is your second team-based assignment. As in the past assignment, you must work in a team of two for this during the
laboratory session and throughout the coming week. You may select your own partner, as long as you are not working with the same
person from the previous assignment.  As this is the most challenging assignment so far this semester, you
and your partner should plan your time this week accordingly and work on it incrementally. Please make sure that you use
Slack and your team's Git repository to collaborate effectively. Remember, divide and conquer!

\vspace{-0.05in}
\subsection*{General Guidelines for Labs}
\vspace{-0.05in}
\begin{itemize}
\item
{\bf Work on the Alden Hall computers.} If you want to work on a different
machine, be sure to transfer your programs to the Alden
machines and re-run them before submitting.
\item
{\bf Update your repository often!} You should add, commit,
and push your updated files each time you work on them.  I will not grade
your programs until the due date has passed.
\item
{\bf Review the Honor Code policy.} You
may discuss programs with others, but programs that are nearly identical
to others will be taken as evidence of violating the Honor Code.
\end{itemize}

\vspace{-0.05in}
\subsection*{Reading Assignment}
\vspace{-0.05in}
To learn more about Java strings and random numbers, review Sections 3.1--3.5 in your textbook. Also carefully study the sample program
described in the section called {\tt A Sample Program} in this lab assignment document.

\vspace{-0.05in}
\subsection*{Create a new directory and a Java program}
\vspace{-0.05in}
In your own {\tt cs111F2014-<your user name>} repository inside {\tt labs} directory, create a directory called {\tt lab5} by typing ``{\tt mkdir lab5}''. Type ``{\tt cd lab5}'' to move to the new directory.  Then using {\tt gvim} create a Java program named {\tt Lab5.java}. See the section called {\tt DNA Manipulation} in this document for the specifics on what you need to do in your {\tt Lab5.java} program.

\vspace{-0.05in}
\subsection*{Study a Sample Program}
\vspace{-0.05in}
Go to the shared course repository and pull {\tt StringDemo.java}. Copy this program into {\tt lab5} directory
inside your own {\tt cs111F2014-<your user name>} repository.
Open this program and examine it to see what it does. Then
run it a few times with different input strings.\\

\noindent{\bf Questions to Discuss with Your Group Member}
(not to be handed in):
\begin{itemize}
\item Where is the random number generator {\em declared}?
\item Suppose you name the random number generator {\tt wilbur} instead of {\tt r}.
What other lines in the program need to be changed?
\item Where is the scanner {\em declared}?
\item Suppose you name the scanner {\tt orville} instead of {\tt scan}.
What other lines in the program need to be changed?
\item
Suppose the line:
\begin{center}
\verb$s1 = s1.toUpperCase();$
\end{center}
is changed to:
\begin{center}
\verb$s1.toUpperCase();$
\end{center}
Will the program still correctly display the upper-case version of {\tt s1}?
\item
Do the statements {\tt s1.toUpperCase();} and {\tt s1.toLowerCase();}
change the contents of {\tt s1}? (See previous question!)
\item
Suppose the user types {\tt abcde} when asked to enter a string.
\begin{itemize}
\item Write this down with numbers to indicate the positions (indices) of each character.
\item
What position number is immediately to the left of the letter {\tt a}?
\item
What position number is immediately to the right of the letter {\tt e}?
\item
How many different ways are there to insert an ``{\tt x}'' into the string
{\tt abcde}? List all the position numbers where this {\tt x} could be placed.
(Don't forget the beginning and end.)
\item
According to the book, the {\tt nextInt(num)} method of the {\tt Random}
class returns a random number in the range 0 to {\tt num - 1}.
List all the values that could be returned by ``{\tt r.nextInt(5)}''.
\item
Answer the question in the comment next to the statement ``\verb$location = r.nextInt(len+1);$''.
\end{itemize}
\item
On paper, write down the string {\tt ABCDEFG}, with position numbers
between the letters.
\begin{itemize}
\item
Underline the portion corresponding to the
expression:
\begin{center}
\verb$"ABCDEFG".substring(0,3)$
\end{center}
\item
Underline the portion corresponding to the
expression:
\begin{center}
\verb$"ABCDEFG".substring(3)$
\end{center}
\item
What string do we get if we evaluate the expression:
\begin{center}
\verb$"ABCDEFG".substring(0,3) + "ABCDEFG".substring(3)$
\end{center}
\end{itemize}
Explain, in English words, what the following statement does:
\begin{center}
\verb$s2 = s1.substring(0,location) + `x' + s1.substring(location);$
\end{center}
\item
What is the value of the expression \verb$"PQRST".charAt(0)$ ?
\item
What is the value of the expression \verb$"PQRST".charAt(2)$ ?
\item
What is the value of the expression \verb$"PQRST".charAt(4)$ ?
\item
List all possible values that can be returned by the expression
\verb$r.nextInt(5)$
\item
Explain, in English words, what the following statement does:
\begin{center}
\verb$c = "PQRST".charAt(r.nextInt(5));$
\end{center}
\end{itemize}
Don't continue with the assignment until you understand the answers to all the
questions above! {\em Please} feel free to discuss these with your
classmates, teaching assistants and your instructor!

\vspace{-0.05in}
\subsection*{DNA Manipulation Program}
\vspace{-0.05in}
Bioinformatics is the study of biological phenomena by the use of biology, mathematics and computer science. One of the most important study areas in bioinformatics concerns DNA. Deoxyribonucleic acid (DNA) is a molecule that encodes the genetic instructions (genes) which are used by all known living organisms and many viruses to build the proteins required to sustain existence. The genes of DNA are written in the nucleotides; guanine (G), adenine (A), thymine (T), and cytosine (C), (chemical compounds) which serve as the alphabet of the genetic language.
Essentially, a DNA string is a string consisting of only the letters {\tt A}, {\tt C}, {\tt G}, and {\tt T}, for instance, ``{\tt CAATGTCAC}''. These strings encode
various genetic traits such as hair color, eye color, and many others.

Each DNA string has a {\em complement} formed by replacing each code letter by its
complementary code. {\tt A} and {\tt T} are complements; so are {\tt G} and
{\tt C}. Thus, the complement to the string ``{\tt CAATGTCAC}'' is
``{\tt GTTACAGTG}''.
DNA sometimes undergoes a {\em mutation}. There are three types of mutation:
insertion of a new letter somewhere in the string; removal of a letter from the
string; and replacement of one letter by another. The table below shows the examples of the replacement of letters and the complement of the given sequence.

\begin{center}

\begin{tabular}{|cc|}
\hline
\textbf{Strand} & \textbf{Sequence}\\
\hline\hline
$S$      & $\mathtt{ACGTGCCTCTTGGTAC}$ \\
\hline
A $\to$ T &  $\mathtt{TCGTGCCTCTTGGTTC}$ \\
T $\to$ A &  $\mathtt{TCGAGCCACAAGGATC}$ \\
C $\to$ G &  $\mathtt{TGGAGGGAGAAGGATG}$ \\
G $\to$ C &  $\mathtt{TGCACGGAGAACCATG}$ \\
\hline
$S_{complementary}$ &  $\mathtt{TGCACGGAGAACCATG}$ \\
\hline
\end{tabular}
\end{center}

Write a Java program that does the following.
Note that all changes are made to the {\em original} input string; they
are not cumulative.
\begin{enumerate}
\item
Ask the user to type in a string of DNA, declare a variable and save the DNA string that the user inputs into a variable called ``{\tt dnaString}''.
\item
Print the complement of {\tt dnaString}, appropriately labeled.
(HINT: use several applications of the String class's ``{\tt replace}'' method. For example, if I want to replace all characters `A' with characters `T' in the String variable {\tt dnaString}, then I will say {\tt dnaString.replace(`A',`T');}, but there's still a trick in this case of getting the complement!)

Another HINT regarding the trick: As you make your substitutions to get your complementary string, remember that you are replacing, for example, \emph{A}'s to \emph{T}'s from the $S$ to make our complementary sequence ($S_{complementary}$). Do not simply perform such substitution directly because you will be unable to figure out which \emph{T}'s were original \emph{T}'s (and not the replaced ones) that you were supposed to change to \emph{A}'s. You can check your complementary sequence from the website  (using the \emph{Complement} option after you enter your sequence): \url{arep.med.harvard.edu/labgc/adnan/projects/Utilities/revcomp.html}.

\item[NOTE:]
The next three parts of your program will also use Java's {\tt Random} class. You may want to read about the {\tt Random} class at the bottom of this document to understand the concepts behind the {\tt Random} class better.
\item
Perform a random mutation consisting of inserting a randomly-chosen
extra letter into {\tt dnaString}; it
must be one of the four allowed letters. Print this, appropriately labeled
and identifying the position of the insertion and the letter inserted.
\item
Perform a random mutation consisting of removing a letter from a
randomly-chosen position in {\tt dnaString}; print this, appropriately labeled
and identifying the position of the insertion and the letter removed.
\item
Perform a random mutation consisting of altering a single letter from a
randomly-chosen position in {\tt dnaString}; it must be changed to a
randomly-chosen letter from the set of allowed letters. Print it, appropriately
labeled
and identifying the position of the replacement, the new letter, and the
letter it replaces.
\end{enumerate}


Here are two sample runs. You will get different values, of course, since the
changes are random, but the structure of the output will be the same:
\begin{verbatim}
     aldenv5:lab5 jjumadinova$ java Lab5
     Janyl Jumadinova
     Lab 5
     Thu Oct 2 12:51:58 EDT 2014

     Enter a string containing only C, G, T, and A: actg
     Complement of ACTG is TGAC
     Inserting T at position 0 gives TACTG
     Deleting from position 1 gives ATG
     Changing position 2 gives ACGG
\end{verbatim}
\begin{verbatim}
     aldenv5:lab5 jjumadinova$ java Lab5
     Janyl Jumadinova
     Lab 5
     Thu Oct 2 12:52:58 EDT 2014

     Enter a string containing only C, G, T, and A: actg
     Complement of ACTG is TGAC
     Inserting G at position 0 gives GACTG
     Deleting from position 0 gives CTG
     Changing position 0 gives ACTG
\end{verbatim}

In the second example, nothing was changed in the last line---the program
randomly replace the letter ``{\tt A}'' with the letter ``{\tt A}'' !
This behavior is ok.

\vspace{-0.05in}
\subsection*{Additional Program Requirements}
\vspace{-0.05in}
\begin{itemize}
\item Make sure your program prints the names of all team members, the lab number, and the date as the first few output lines.
\item Make sure your program contains the comment header with the honor pledge, the names of all team members, lab number, date, and the purpose of the program.
\item Make sure you document your program properly, by using comments throughout your program whenever appropriate.
\item Make sure your output
is neat (no missing spaces, etc.) and that your program is neat (indenting, etc.).
\end{itemize}

\vspace{-0.05in}
\subsection*{Required Deliverables}
\vspace{-0.05in}
In addition to turning in printed and signed versions, for this assignment you are invited to submit electronic versions of the following deliverables through the Bitbucket repository. As you complete this step, you should make sure that you
created a {\tt lab5/} directory within the Git repository.  Then, you can save all of the required deliverables in the
{\tt lab5/} directory---please see the course instructor or a teaching assistant if you are not able to create your
directory properly.

\begin{enumerate}
	\item A completed, properly commented and formatted {\tt Lab5.java} program.

        \item An output document containing at least \textbf{three different} outputs obtained after running {\tt Lab5} in the terminal window at least three times.

        \item A document describing the strategy your team developed for completing this assignment and the work each team member has completed. While the {\tt Lab5} program and the output file have to be the same for both members of the same team, you may include any challenges you individually faced in this document.
\end{enumerate}

\vspace{-0.1in}
Share your program, the output file and the document describing your team work with me through your Git repository by correctly using ``{\tt git add}'', ``{\tt
git commit}'', and ``{\tt git push}'' commands. When you are done, please ensure that the Bitbucket Web site has
a {\tt lab5/} directory in your repository with the two files called {\tt Lab5.java}, {\tt output} and {\tt report}. You should see
the instructor if you have questions about assignment submission.

\vspace{-0.05in}
\subsection*{A Quick Tutorial on Random Number Generation}
\vspace{-0.05in}
To generate random numbers, we need an object of the {\tt Random} class.
You can name this variable anything you want---``{\tt rand}'' or ``{\tt
random}'' are good names. To create a new random number generator
named {\tt rand}, be sure to import the {\tt Random} class:
\begin{center}
\verb$ import java.util.Random;$
\end{center}
and then create an instance of this class:
\begin{center}
\verb$Random rand = new Random();$
\end{center}

The three most useful methods in the {\tt Random} class are {\tt nextInt}, {\tt nextFloat} and {\tt nextDouble}. If {\tt rand} is the name of our random number
generator (it can be called something else) and {\tt n} is a positive integer, {\tt rand.nextInt(n)}
produces an {\tt int} in the range 0, \ldots, {\tt n}-1 and {\tt
rand.nextDouble()} and {\tt rand.nextFloat()} produce a {\tt double} and a {\tt float} respectively in the range from 0 to 1 (not
including 1).

By being clever, we can get different ranges---here are a few
examples. In the Java code, assume {\tt i} is an {\tt int} variable,
{\tt d} is a {\tt double} variable, and {\tt rand} is an object of the
{\tt Random} class:
\begin{center}
\begin{tabular}{p{2.5in}p{3.5in}}
\multicolumn{1}{c}{\bf Desired Range} & \multicolumn{1}{c}{\bf Java
Statement}\\\hline
0, 1, 2, 3, 4, 5 & \verb$i = rand.nextInt(6);$\\
10, 11, \ldots, 19, 20 & \verb$i = rand.nextInt(11) + 10;$\\
-5, -4, -3, \ldots, 4, 5 & \verb$i = rand.nextInt(11) - 5;$\\
0, 3, 6, 9, 12 & \verb$i = 3 * rand.nextInt(5);$\\
-1, 1 & \verb$i = 2 * rand.nextInt(2) - 1;$\\
$0 \leq d < 1$ & \verb$d = rand.nextDouble();$\\
$0 \leq d < 10$ & \verb$d = 10 * rand.nextDouble();$\\
$-5 \leq d < 5$ & \verb$d = 10 * rand.nextDouble() - 5;$\\
0.0, 0.1, 0.2, \ldots, 0.9, 1.0 & \verb$d = rand.nextInt(11)/10.0;$
\end{tabular}
\end{center}

You can also use random numbers to do other creative things, for example pick out characters randomly. Let's say I want
to pick out a random letter out of a first name, I can first create and initialize a variable {\tt{String professorName =
"Janyl";}} Then I can generate a random number between $0$ and $4$ and save it into a variable of type {\tt int} as
{\tt{ int randomNumber = rand.nextInt(5)}}. Finally, I  use {\tt{charAt}} method to select the character at a position
created by the random number generator as {\tt{professorName.charAt(randomNumber)}}.

\end{document}
