\input{111pre}
\begin{document}
\MYTITLE{Quiz 1 Study Guide \\ Delivered: Monday, September 21, 2015 \\ Quiz: Friday, September 25, 2015, 11:00 am}

This course will have its first quiz on Friday, September 25, 2015 from 11:00 to 11:50 am. The quiz will be ``closed
notes'' and ``closed book'' and it will cover the following materials. Please review the ``Course Schedule'' on the Web
site for the course to see the content and slides that we have covered to this date. Students may post questions about
this material to our Slack channel.

\begin{itemize}

  \item Chapter One in Lewis \& Loftus (e.g., ``Introduction to Computation and Programming'')

  \item Chapter Two in Lewis \& Loftus, Sections 2.1--2.9 (e.g., ``Data and Expressions'')

  \item Using the basic commands in the Linux operating system; editing in {\tt gvim}, compiling and executing
    programs in Linux; knowledge of the basic commands for using {\tt git} and Bitbucket.

  \item Your class notes and lecture slides, labs 1--4, practicals 1--3, and the handouts from lab.

\end{itemize}

\noindent The quiz will be a mix of questions that have a form such as fill in the blank, short answer, true/false, and
completion.  The emphasis will be on the following topics:

\begin{itemize}

\item Fundamental concepts in computing and the Java language (e.g., definitions and background)

\item Practical laboratory techniques (e.g., editing, compiling, and running programs; effectively using files and
  directories; correctly using Bitbucket through the command-line {\tt git} program)

\item Understanding Java programs (e.g., given a short, perhaps even one line, source code segment written in Java,
  understand what it does and be able to precisely describe its output).

\item Composing Java statements and programs, given a description of what should be done. Students should be completely
  comfortable writing short source code statements that are in nearly-correct form as Java code. While your program may
  contain small syntactic errors, it is not acceptable to ``make up'' features of the Java programming language that do
  not exist in the language itself---so, please do not call a ``{\tt solveQuestionThree()}'' method!

\end{itemize}

\noindent No partial credit will be given for questions that are true/false, completion, or fill in the blank. Minimal
partial credit may be awarded for the questions that require a student to write a short answer. You are strongly
encouraged to write short, precise, and correct responses to all of the questions. When you are taking the quiz, you
should do so as a ``point maximizer'' who answers first the questions that you are most likely to answer correctly for
full points. Please keep the time limitation in mind as you are absolutely required to submit the examination at the end
of the class period unless you have written permission for extra time from a member of the Learning Commons. Students
who do not submit their quiz on time will have their overall point total reduced.

\noindent Students are required to fully adhere to the Honor Code during the completion of this quiz. More details about
the Allegheny College Honor Code are provided on the syllabus. Students are strongly encouraged to carefully review the
full statement of the Honor Code before taking this quiz.



\end{document}
